Testing atlantisom: do we ever get small fish in length comps?

### 2019-05-15 TestSmallestLength.Rmd
looking for small turbot

---
title: "Testing atlantisom: generate census length comps for Norwegian Barents Sea Atlantis"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

@@

## Simulate a survey part 4: sample for length composition

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages:

```{r census-spec, message=FALSE, warning=FALSE, echo=TRUE}

@@

# not enough memory to do all NOBA outputsteps, change to once per year
midptyr <- round(median(seq(1,stepperyr)))

annualmidyear <- seq(midptyr, noutsteps, stepperyr)

@@

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition. *Sampling once per year, rather than output timestep.*

```{r stdsurveyNbased, echo=TRUE}

survey_testNall <- create_survey(dat = truth$nums,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# consider saving this interim step if it takes a long time go generate

@@

```{r truecohortagecomp}
# what is true composition? need annual by species, use code from sample_fish
# do tidyly
dat2 <- survey_testNall %>%
  group_by(species, agecl, time) %>%
  summarize(numAtAge = sum(atoutput))

totN <- dat2 %>%
  group_by(species, time) %>%
  summarize(totN = sum(numAtAge))

dat2totN <- merge(dat2, totN)

# ageclcomp <- ggplot(dat2totN, aes(x=agecl, y=numAtAge/totN, col=time)) +
#   geom_point()
# 
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 1, scales="free")
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 2, scales="free")
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 3, scales="free")
# ageclcomp + facet_wrap_paginate(~species, ncol=3, nrow = 3, page = 4, scales="free")

@@

Then we use the `sample_fish` function with very high effN and compare to true annual age comp calculated above as a test, which matches as we have shown [here](https://sgaichas.github.io/poseidon-dev/TrueLengthAgeCompTest.html).

The following gets both numbers and weights to get the length comp, and ultimately the weight at age for use in assessments.

First we subset the nums, resn, and structn components of `run_truth` using the same `create_survey` design selecting boxes, time (*once per year for NOBA*), and species. We need only apply survey efficiency and selectivity to nums, however (done above), so we apply `aggregateDensityData` here.

```{r aggdens, echo=TRUE}

# aggregate true resn per survey design
survey_aggresnall <- aggregateDensityData(dat = truth$resn,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxall)

# aggregate true structn per survey design
survey_aggstructnall <- aggregateDensityData(dat = truth$structn,
                                 time = annualmidyear,
                                 species = survspp,
                                 boxes = boxall)

Now we should have inputs to `sample_fish` on the same scale, and they need to be aggregated across boxes into a single biological sample for the whole survey. We are not maintaining spatial structure in sampling because it isn't used in most assessments.

To do the proper aggregation and not apply the multinomial sampling to the density data, I rewrote `sample_fish` to apply the median if `sample=FALSE` in the function call.

```{r censussurvey-samplefish, warning=FALSE, message=FALSE, echo=TRUE}

# apply default sample fish as before to get numbers
numsallhigh <- sample_fish(survey_testNall, effNhigh)

#dont sample these, just aggregate them using median
structnall <- sample_fish(survey_aggstructnall, effNhigh, sample = FALSE)

resnall <-  sample_fish(survey_aggresnall, effNhigh, sample = FALSE)

Select a single species for testing. Generating length composition for a single species should take just a few minutes to run. Update: for NOBA with many output timesteps, even a single species took 15-20 minutes.

```{r singlespp-lengthcomp, echo=TRUE}

# this should still represent a census but with polygon and layer aggregated
# WARNING: selecting a species by name is hardcoded for a particular model

ss_numsallhigh <- numsallhigh[numsallhigh$species == "Green_halibut",]
ss_structnall <- structnall[structnall$species == "Green_halibut",]
ss_resnall <- resnall[resnall$species == "Green_halibut",]

ss_length_censussurvsamp <- calc_age2length(structn = ss_structnall,
                                 resn = ss_resnall,
                                 nums = ss_numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

This gives annual length comps:

```{r ss-lengthsamp1-test}

lfplot <- ggplot(ss_length_censussurvsamp$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

# try dir = "v" option for vertical lf comparisons

## Run for all species then SAVE

This code block can take hours (5+ for CCA and estimated 17+ for NOBA, but it never finished after 32+) to run, so be prepared, but then the saved file can be read back in to have length comps for all surveyed species. For a normal survey query we would have only one output timestep per year, which should save some time.

And lets change NOBA to estimate length only 1 timestep per year (~2.5 hours to run):

```{r allspplength, echo=TRUE, eval=FALSE}

length_censussurvsamp <- calc_age2length(structn = structnall,
                                 resn = resnall,
                                 nums = numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

#save for later use, takes a long time to generate
saveRDS(length_censussurvsamp, file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))

Read it back in:

```{r loadlengthcomp, echo=TRUE}

length_censussurvsamp <- readRDS(file.path(d.name, paste0(scenario.name, "length_censussurvsamp.rds")))

Demo plots for other species:

```{r moreplots, echo=TRUE}

ss_length_censussurvsamp <- length_censussurvsamp$natlength %>%
  filter(species == "Norwegian_ssh")

lfplot <- ggplot(ss_length_censussurvsamp, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte()

@@

So this looks like our best approximation of a true length comp for a particular Atlantis run (although it seems critical to limit output to allow functions to finish). Comparisons with standard surveys next.

>>>>

---
title: "Testing atlantisom: do we ever get small fish in length comps?"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
bibliography: "packages.bib"
---

@@

Here we further test the function `calc_age2length` to see if small fish ever appear with full census sampling. If they don't, perhaps the default CV in the function needs modification? This question is triggered by the [sampling for NOBA Greenland halibut](https://sgaichas.github.io/poseidon-dev/NOBAStdSurvLengthCompTest.html), in which fish below 50 cm didn't show up.

For all setup, etc, please see previous files Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/TrueBioTest.html) and  [here](https://sgaichas.github.io/poseidon-dev/TrueLengthCompTest.html).

@@

## Simulate a survey part 4: sample for length composition (testing revised function)

Full methods are explained [here](https://sgaichas.github.io/poseidon-dev/StdSurvLengthCompTest.html).

We will apply examples here to only one species, Greenland halibut in NOBA, which grows to a large size.

To create a census, the user specifies the timing of the survey, which species are captured, the spatial coverage of the survey, the species-specific survey efficiency ("q"), and the selectivity at age for each species. The following settings should achieve a survey that samples all Atlantis model output timesteps, all fish and shark species, and all model polygons, with perfect efficiency and full selectivity for all ages.

```{r census-spec, message=FALSE, warning=FALSE}

@@

@@

Here we use `create_survey` on the numbers output of `run_truth` to create the survey census of age composition (for just one species in this case). The `sample_fish` applies the median for aggregation and does not apply multinomial sampling if `sample=FALSE` in the function call.

Because we don't want to wait 24 hours for this, we will look at only the first 112 time steps.

```{r stdsurveyNbased-GHR, echo=TRUE}

ss.name <- funct.group.names[funct.group.names == "Green_halibut"]

# get survey nums with full (no) selectivity
ss_survey_testNall <- create_survey(dat = truth$nums,
                                 time = c(0:111),
                                 species = ss.name,
                                 boxes = boxall,
                                 effic = effic1,
                                 selex = selex1)

# apply default sample fish as before to get numbers
ss_numsallhigh <- sample_fish(ss_survey_testNall, effNhigh)

# aggregate true resn per survey design
ss_aggresnall <- aggregateDensityData(dat = truth$resn,
                                 time = c(0:111),
                                 species = ss.name,
                                 boxes = boxall)

# aggregate true structn per survey design
ss_aggstructnall <- aggregateDensityData(dat = truth$structn,
                                 time = c(0:111),
                                 species = ss.name,
                                 boxes = boxall)

#dont sample these, just aggregate them using median
ss_structnall <- sample_fish(ss_aggstructnall, effNhigh, sample = FALSE)

ss_resnall <-  sample_fish(ss_aggresnall, effNhigh, sample = FALSE)

@@

Length sample with user specified max length bin (200 cm):

```{r userset-maxlen, echo=TRUE}

ss_length_census <- calc_age2length(structn = ss_structnall,
                                 resn = ss_resnall,
                                 nums = ss_numsallhigh,
                                 biolprm = truth$biolprm, fgs = truth$fgs,
                                 maxbin = 200,
                                 CVlenage = 0.1, remove.zeroes=TRUE)

@@

We should get the upper end of Greenland halibut with a 200cm max length bin. What I am concerned about here is whether we get very small fish at any time step:

```{r sslengthsamp2-test}

lfplot <- ggplot(ss_length_census$natlength, aes(upper.bins)) +
  geom_bar(aes(weight = atoutput)) +
  theme_tufte() +
  labs(subtitle = paste(scenario.name, ss_length_census$natlength$species))

@@

Are there fish <50 cm long at any output timestep? If not, does this make sense? There appear to be a few, such as timestep 4. Hard to see at this level.

Aggregate length comp over all of these time steps (first 22+ years):

```{r agglengthcomp}

lfplot

This suggests that the smallest Greenland turbot in the ecosystem are ~25 cm over all of the timesteps we checked. Does this make sense?

Or do we need to change something about the length comp estimation?
